From Reversible Programs to \\ Univalent Universes and Back

### 2021-07-04 Pi+/paper/reversible1.tex
Cut more in section 3
proof outline for thm1

\section{A Reversible Programming Language: Set-Theoretic Semantics}
%% ~3 pages
\label{sec:reversibleone}
\label{sec:reversibletwo}
\label{langeqeq}
\label{sec:informal}

@@

\noindent Like the original circuit, we examine the bit at index 1 (corresponding to the component $B$ in a tuple
$(A,(B,C))$): if the bit is true, we perform an \verb|x| operation on component $A$, and otherwise we perform a
\verb|cx| operation on $C,A$. The two uses of \verb|x| gates in the circuit are now unnecessary as they were only needed
to encode a two-way conditional expression using a sequence of one-way conditional expressions (the only ones available in
the linear circuit model).

@@

\begin{theorem}\label{thm:semone}
  The denotational semantics is sound in the following sense:
  \begin{itemize}
    \item For every level-1 combinator $c : A \isoone B$, we have that $\denot{c}$ is a bijection between $\denot{A}$ and $\denot{B}$.
      \item For every pair of combinators $c_1$ and $c_2$ of the same type $A \isoone B$, if there exists a level-2
        combinator $\alpha$ such that $\alpha : c_1 \isotwo c_2$, then $\denot{c_1} = \denot{c_2}$ using
        extensional equivalence of functions.
  \end{itemize}
\end{theorem}
%% proof?

>>>>

\section{A Reversible Programming Language}
%% ~3 pages
\label{sec:reversibleone}
\label{sec:reversibletwo}
\label{langeqeq}
\label{sec:informal}

@@

\noindent Like the original circuit, we examine the bit at index 1 (corresponding to the component $B$ in a tuple
$(A,(B,C))$): if the bit is true, we perform an \verb|x| operation on component $A$, and otherwise we perform a
\verb|cx| operation on $(C,A)$. The two uses of \verb|x| gates in the circuit are now unnecessary as they were only needed
to encode a two-way conditional expression using a sequence of one-way conditional expressions (the only ones available in
the linear circuit model).

@@

\begin{theorem}\label{thm:semone}
  The denotational semantics is sound in the following sense:
  \begin{itemize}
    \item For every level-1 combinator $c : A \isoone B$, we have that $\denot{c}$ is a bijection between $\denot{A}$ and $\denot{B}$.
    \item For every pair of combinators $c_1$ and $c_2$ of the same type $A \isoone B$, if there exists a level-2
        combinator $\alpha$ such that $\alpha : c_1 \isotwo c_2$, then $\denot{c_1} = \denot{c_2}$ using
        extensional equivalence of functions.
  \end{itemize}
\end{theorem}
\begin{proof}
  For every primitive combinator $c$ listed on one side of Fig.~\ref{fig:pi-terms}, let $!c$ be the combinator listed on
  the other side. Thus $! \assoclp$ is $\assocrp$ and $! \swapp$ is $\swapp$ itself. Then we have that $\denot{c}$ and
  $\denot{!c}$ is an equivalence. For the level-2 combinator \Afun{idr◎l}, we check
  $\denot{\AgdaBound{c}~\AgdaOperator{\AgdaInductiveConstructor{◎}}~\AgdaInductiveConstructor{id⟷₁}}
  = \mathit{id} \circ \denot{c} = \denot{c}$. The other cases are only slightly more involved calculations.
\end{proof}
